Why is it that statistical tests for residuals are not widely used? An explanation using visual inference.

### 2023-03-01 paper_comparison/paper_comparison/paper_comparison.Rmd
update the power cmp part

Effect size $E$ can be defined as the difference of a parameter for a particular model or distribution, or a statistic derived from a sample. Importantly, it needs to reflect the treatment we try to measure. According to the underlying simulated model, we define the effect size $E = \sqrt{n}/(\sigma k)$ for the non-linearity model and $E = \sqrt{n}/(|a| + 1)$ for the heteroskedasticity model. The parameter $k$ is the number of turning points embedded in the residual plot. It is used as a penalty term to reduce the effect size for complex shapes. There are one, two, three and five turning points in "U", "S", "M", "Triple-U" shape respectively.  The denominator of the effect size of the heteroskedasticity model is $|a| + 1$. It is used to

@@

To compare the power of visual tests and conventional tests, we simulate 200000 samples evenly from the models used in the human subject experiments, and record the conventional $p$-values produced from every sample. We pick RESET test (`resettest`) and Breusch-Pagan test (`bptest`) from the R package `lmtest`, and Shapiro-Wilk test (`shapiro.test`) from the built-in R package `stats`. Among them, RESET test is the only exact and appropriate test in this scenario. Both the Breusch-Pagan test and the Shapiro-Wilk test are approximate and inappropriate tests. Their estimated power is shown in Figure \ref{fig:polypower}. To set up the RESET test, we include different powers of fitted values as proxies. According to @ramsey_tests_1969, there are no general rules for the power of the fitted values needed by the RESET test, but it finds power up to four is usually sufficient. Thus, we follow this guideline to conduct the RESET test. For the Breusch-Pagan test, the choice of regressors in the auxiliary regression is left to the user [@breusch_simple_1979]. But as @waldman1983note suggested, it is a good choice for the set of auxiliary regressors in the Breusch-Pagan test be the same as the White test. Thus, we include both $\boldsymbol{x}$ and $\boldsymbol{x}^2$ in the auxiliary regression.

@@

target_sigma <- c(3.65, 2.84, 2.21, 1.72, 1.34, 1.04, 0.81, 0.63, 0.49, 0.38, 0.3)

# Calculate the effect size 
es <- map_dbl(target_sigma, 
              ~log(poly_model(shape = 2, 
                              x = {raw_x <- rand_normal(sigma = 0.3); closed_form(~stand_dist(raw_x))}, 
                              sigma = .x)$
                     average_effect_size(n = 100, type = "simple")
                   )
              )
```

@@

# Use area() to design layout
map(1:11, ~patchwork::area(21, .x, 22, .x)) %>%
  reduce(c) %>%
  c(patchwork::area(3, 1, 20, 11), .) %>%
  patchwork::wrap_plots(append(plist, list(p), after = 0), 
                        design = .)
```

@@

  # Theme
  theme_light(base_size = 5) +
  theme(panel.grid.major.x =  element_line(colour = "grey70", linewidth = 0.5)) +
  scale_color_manual(values = rev(rcartocolor::carto_pal(4, "Vivid"))) +
  scale_x_continuous(breaks = seq(1, 7, 1), minor_breaks = NULL) +
  xlab(expression(log[e] (Effect_size))) +
  ylab("Power") +
  labs(col = "", size = "# lineups")
```

@@

target_b <- c(0.17, 0.23, 0.32, 0.46, 0.67, 1.04, 1.7, 3, 6.3, 16.2, 61)

# Calculate the effect size 
es <- map_dbl(target_b, 
              ~log(heter_model(a = 1, 
                               b = .x, 
                               x = {raw_x <- rand_normal(sigma = 0.3); closed_form(~stand_dist(raw_x))})$
                     average_effect_size(n = 100, type = "simple")
                   )
              )
```

>>>>

(do we actually need the penalty term?)

Effect size $E$ can be defined as the difference of a parameter for a particular model or distribution, or a statistic derived from a sample. Importantly, it needs to reflect the treatment we try to measure. According to the underlying simulated model, we define the effect size $E = \sqrt{n}/(\sigma k)$ for the non-linearity model and $E = b\sqrt{n}/(|a| + 1)$ for the heteroskedasticity model. The parameter $k$ is the number of turning points embedded in the residual plot. It is used as a penalty term to reduce the effect size for complex shapes. There are one, two, three and five turning points in "U", "S", "M", "Triple-U" shape respectively.  The denominator of the effect size of the heteroskedasticity model is $|a| + 1$. It is used to

@@

We pick RESET test (`resettest`) and Breusch-Pagan test (`bptest`) from the R package `lmtest`, and Shapiro-Wilk test (`shapiro.test`) from the built-in R package `stats`. Among them, RESET test is the only exact and appropriate test in this scenario. Both the Breusch-Pagan test and the Shapiro-Wilk test are approximate and inappropriate tests. Their estimated power is shown in Figure \ref{fig:polypower}. To set up the RESET test, we include different powers of fitted values as proxies. According to @ramsey_tests_1969, there are no general rules for the power of the fitted values needed by the RESET test, but it finds power up to four is usually sufficient. Thus, we follow this guideline to conduct the RESET test. For the Breusch-Pagan test, the choice of regressors in the auxiliary regression is left to the user [@breusch_simple_1979]. But as @waldman1983note suggested, it is a good choice for the set of auxiliary regressors in the Breusch-Pagan test be the same as the White test. Thus, we include both $\boldsymbol{x}$ and $\boldsymbol{x}^2$ in the auxiliary regression.

@@

target_sigma <- c(5, 3.03, 1.84, 1.12, 0.68, 0.41, 0.25)

# Calculate the effect size 
es <- map_dbl(target_sigma, 
              function(.x){
                mod <- poly_model(shape = 2, 
                              x = {raw_x <- rand_normal(sigma = 0.3); closed_form(~stand_dist(raw_x))}, 
                              sigma = .x)
                log(mod$sample_effect_size(mod$gen(100), type = "simple"))
                }
              )
```

@@

# Use area() to design layout
map(1:7, ~patchwork::area(21, .x, 22, .x)) %>%
  reduce(c) %>%
  c(patchwork::area(3, 1, 20, 7), .) %>%
  patchwork::wrap_plots(append(plist, list(p), after = 0), 
                        design = .)
```

@@

  # Theme
  theme_light(base_size = 5) +
  theme(panel.grid.major.x =  element_line(colour = "grey70", linewidth = 0.5)) +
  scale_color_manual(values = rev(rcartocolor::carto_pal(4, "Vivid"))) +
  # scale_size_manual(values = seq(1, 9)) +
  scale_x_continuous(breaks = seq(1, 7, 1), minor_breaks = NULL) +
  xlab(expression(log[e] (Effect_size))) +
  ylab("Power") +
  labs(col = "", size = "# lineups")
```

@@

target_b <- c(0.54, 1.48, 4.02, 10.92, 29.68, 80.69, 219.33)

# Calculate the effect size 
es <- map_dbl(target_b, 
              function(.x){
                mod <- heter_model(a = 1, 
                                   b = .x, 
                                   x = {raw_x <- rand_normal(sigma = 0.3); closed_form(~stand_dist(raw_x))})
                log(mod$sample_effect_size(mod$gen(100), type = "simple"))
                }
              )
```

### 2023-03-01 paper_comparison/paper_comparison/paper_comparison.tex
update the power cmp part

\hypertarget{caculation-of-p-value-and-power}{%
\section{Caculation of p-value and
power}\label{caculation-of-p-value-and-power}}

\hypertarget{p-value}{%
\subsection{\texorpdfstring{\(p\)-value}{p-value}}\label{p-value}}

In hypothesis testing, a \(p\)-value is defined as the probability of
observing test results as least as extreme as the observed result given
\(H_0\) is true. Within the context of visual inference, by involving
\(k\) independent observers, the \(p\)-value can be interpreted as the
probability of having as many or more subjects detect the data plot than
the observed result.

@@

--\textgreater{}

@@

@@

\hypertarget{effect-size}{%
\subsection{Effect size}\label{effect-size}}

@@

@@

To compare the power of visual tests and conventional tests, we simulate
200000 samples evenly from the models used in the human subject
experiments, and record the conventional \(p\)-values produced from
every sample. We pick RESET test (\texttt{resettest}) and Breusch-Pagan
test (\texttt{bptest}) from the R package \texttt{lmtest}, and
Shapiro-Wilk test (\texttt{shapiro.test}) from the built-in R package
\texttt{stats}. Among them, RESET test is the only exact and appropriate
test in this scenario. Both the Breusch-Pagan test and the Shapiro-Wilk
test are approximate and inappropriate tests. Their estimated power is
shown in Figure \ref{fig:polypower}. To set up the RESET test, we
include different powers of fitted values as proxies. According to
\citet{ramsey_tests_1969}, there are no general rules for the power of
the fitted values needed by the RESET test, but it finds power up to
four is usually sufficient. Thus, we follow this guideline to conduct
the RESET test. For the Breusch-Pagan test, the choice of regressors in
the auxiliary regression is left to the user
\citep{breusch_simple_1979}. But as \citet{waldman1983note} suggested,
it is a good choice for the set of auxiliary regressors in the
Breusch-Pagan test be the same as the White test. Thus, we include both
\(\boldsymbol{x}\) and \(\boldsymbol{x}^2\) in the auxiliary regression.

>>>>

\hypertarget{estimation-of-p-value-and-power}{%
\section{Estimation of p-value and
power}\label{estimation-of-p-value-and-power}}

\hypertarget{p-value}{%
\subsection{p-value}\label{p-value}}

In hypothesis testing, a \(p\)-value is defined as the probability of
observing test results as least as extreme as the observed result given
\(H_0\) is true. Conventional hypothesis tests usually have an existing
method to derive or compute \(p\)-value based on the null distribution.
What we need to discuss in the following is the method to estimate
\(p\)-value for a visual test.

Within the context of visual inference, by involving \(k\) independent
observers, the visual \(p\)-value can be interpreted as the probability
of having as many or more subjects detect the data plot than the
observed result.

@@

As pointed out by \citet{vanderplas2021statistical}, this basic binomial
model doesn't take into account the possible dependencies in the visual
test due to repeated evaluations of the same lineup. And it is
inapplicable to visual test where subjects are asked to select one or
more ``most different'' plots from the lineup.
\citet{vanderplas2021statistical} summarises three common scenarios in
visual inference: (1) \(K\) different lineups are shown to \(K\)
subjects, (2) \(K\) lineups with different null plots but the same data
plot are shown to \(K\) subjects, and (3) the same lineup is shown to
\(K\) subjects. Out of these three scenarios, Scenario 3 is the most
common in previous studies as it puts the least constraints on the
experiment design. For Scenario 3, \citet{vanderplas2021statistical}
models the probability of a plot \(i\) being selected from a lineup as
\(\theta_i\), where \(\theta_i \sim Dirichlet(\alpha)\) for
\(i=1,...,m\) and \(\alpha > 0\). The number of times plot \(i\) being
selected in \(K\) evaluations is denoted as \(c_i\). In case subject
\(j\) makes multiple selections, \(1/s_j\) will be added to \(c_i\)
instead of one, where \(s_j\) is the number of plots subject \(j\)
selected for \(j=1,...K\). This ensures \(\sum_{i}c_i=K\). Since we are
only interested in the selections of the data plot \(i\), the marginal
model can be simplified to a beta-binomial model and thus the visual
p-value is given as

\begin{equation} \label{eq:pvalue-beta-binomial}
P(C \geq c_i) = \sum_{x=c_i}^{K}{K \choose x}\frac{B(x + \alpha, K - x + (m - 1)\alpha)}{B(\alpha, (m-1)\alpha)},\quad \text{for} \quad c_i \in \mathbb{Z}_0^+
\end{equation}

\noindent where \(B(.)\) is the beta function defined as

\begin{equation} \label{eq:betafunction}
B(a, b) = \int_{0}^{1}t^{\alpha - 1}(1-t)^{b-1}dt,\quad \text{where}\quad a,b>0.
\end{equation}

Note that Equation \ref{eq:pvalue-beta-binomial} given in
\citet{vanderplas2021statistical} only works with non-negative integer
\(c_i\). We extend the equation to non-negative real number \(c_i\) by
applying a linear approximation

\begin{equation} \label{eq:pvalue-beta-binomial-approx}
P(C \geq c_i) = P(C \geq \lceil c_i \rceil) + (\lceil c_i \rceil - c_i) P(C = \lfloor c_i \rfloor), \quad \text{for}\quad c_i \in \mathbb{R}_0^+,
\end{equation}

where \(P(C \geq \lceil c_i \rceil)\) is calculated using Equation
\ref{eq:pvalue-beta-binomial} and \(P(C = \lfloor c_i \rfloor)\) is
calculated by

\begin{equation} \label{eq:pmf-beta-binomial}
P(C = c_i) = {K \choose c_i}\frac{B(c_i + \alpha, K - c_i + (m - 1)\alpha)}{B(\alpha, (m-1)\alpha)},\quad \text{for} \quad c_i \in \mathbb{Z}_0^+.
\end{equation}

Besides, the parameter \(\alpha\) used in Equation
\ref{eq:pvalue-beta-binomial} and \ref{eq:pmf-beta-binomial} is usually
unknown and hence needs to be estimated from the survey data. For low
values of \(\alpha\), only a few plots are attractive to the observers
and tend to be selected. For higher values of \(\alpha\), the
distribution of the probability of each plot being selected is more
even. \citet{vanderplas2021statistical} defines that a plot is
\(c\)-interesting if \(c\) or more participants select the plot as the
most different. Given the definition, The expected number of plots
selected at least \(c\) times, \(E[Z_c]\), is calculated as

\begin{equation} \label{eq:c-interesting-expectation}
E[Z_c(\alpha)] = \frac{m}{B(\alpha, (m-1)\alpha)}\sum_{\lceil c \rceil}^{K}{K \choose x} B(x + \alpha, K - x + (m-1)\alpha).\end{equation}

With Equation \ref{eq:c-interesting-expectation}, \(\alpha\) can be
estimated using maximum likelihood estimation. But for precise estimate
of \(\alpha\), additional human responses to Rorschach lineups, which is
a type of lineup that consists of plots constructed from the same null
data generating mechanism, are required.

@@

The power of a model misspecification test is the probability that the
null hypothesis is rejected given the regression model is misspecified
in a specific way. It is an important indicator when one is concerned
about whether model assumptions have been violated. Although in
practice, one might be more interested in knowing how much the residuals
deviate from the model assumptions, and whether this deviation is of
practical significance.

The power of a conventional hypothesis test is affected by both the true
parameter \(\boldsymbol{\theta}\) and the sample size \(n\). These two
can be quantified in terms of effect size \(E\) to measure the strength
of the residual departures from the model assumptions. Details about the
effect size is provided in Section \ref{effect-size} after the
introduction of the simulation model used in our human subject
experiment. The theoretical power of a test is sometimes not a trival
solution. For a given regression model under a specific effect size, if
the data generating process is also known, the power of a conventional
test can be estimated by counting the number of times the test rejects
the null hypothesis in a repeated sampling exercise and divided by the
number of times simulated from the known data generating process. This
is the method we used to estimate the power of conventional tests in our
experiment.

The power of a visual test on the other hand, may additionally depend on
the ability of the particular subject, as the skill of the individual
may affect the number of observers who identify the data plot from the
lineup \citep{majumder_validation_2013}. To address this issue,
\citet{majumder_validation_2013} models the probability of a subject
\(j\) correctly picking the data plot from a lineup \(l\) using a
mixed-effect logistic regression, with subjects treated as random
effect. Then, the estimated power of a visual test evaluated by a single
subject is the predicted value obtained from the mix-effect model.

However, this mix-effect model does not work with scenario where
subjects are asked to select one or more ``most different'' plots. In
this scenario, having the probability of a subject \(j\) correctly
picking the data plot from a lineup \(l\) is insufficient to determine
the power of a visual test because it does not provide information about
the number of selections made by the subject for the calculation of the
p-value (See Equation \ref{eq:pvalue-beta-binomial-approx}).

Therefore, we directly estimate the probability of a lineup being
rejected by assuming that individual skill has negligible effect on the
variation of the power. This assumption is not necessary true, but it
helps simplifying the model structure, thereby obviate a costly
large-scale experiment to estimate complex covariance matrices. There is
a discussion provided in the Appendix about a potential solution to
model individual effect in the multiple selections scenario.

(should we just link the power together for conventional test?)

We use the same logistic regression model to fit the power of
conventional tests and visual tests formulated as

\begin{equation} \label{eq:logistic-regression-1-1}
Pr(\text{reject}~H_0|H_1,E) = \Lambda\left(log\left(\frac{0.05}{0.95}\right) + \beta_1 E\right),
\end{equation}

\noindent where \(\Lambda(.)\) is the standard logistic function given
as \(\Lambda(z) = exp(z)/(1+exp(z))\). The effect size \(E\) is the only
regressor and the intercept is fixed to \(log(0.05/0.95)\) so that
\(\hat{Pr}(\text{reject}~H_0|H_1,E = 0) = 0.05\).

To study various factors contributing to the power of test, the same
logistic regression model is fit on different subsets of the collated
data grouped by levels of factors. These include the distribution of the
regressor, type of the simulation model and the shape of the residual
departures.

@@

@@

\hypertarget{effect-size}{%
\subsubsection{Effect size}\label{effect-size}}

(do we actually need the penalty term?)

Effect size \(E\) can be defined as the difference of a parameter for a
particular model or distribution, or a statistic derived from a sample.
Importantly, it needs to reflect the treatment we try to measure.
According to the underlying simulated model, we define the effect size
\(E = \sqrt{n}/(\sigma k)\) for the non-linearity model and
\(E = b\sqrt{n}/(|a| + 1)\) for the heteroskedasticity model. The
parameter \(k\) is the number of turning points embedded in the residual
plot. It is used as a penalty term to reduce the effect size for complex
shapes. There are one, two, three and five turning points in ``U'',
``S'', ``M'', ``Triple-U'' shape respectively. The denominator of the
effect size of the heteroskedasticity model is \(|a| + 1\). It is used
to

@@

We pick RESET test (\texttt{resettest}) and Breusch-Pagan test
(\texttt{bptest}) from the R package \texttt{lmtest}, and Shapiro-Wilk
test (\texttt{shapiro.test}) from the built-in R package \texttt{stats}.
Among them, RESET test is the only exact and appropriate test in this
scenario. Both the Breusch-Pagan test and the Shapiro-Wilk test are
approximate and inappropriate tests. Their estimated power is shown in
Figure \ref{fig:polypower}. To set up the RESET test, we include
different powers of fitted values as proxies. According to
\citet{ramsey_tests_1969}, there are no general rules for the power of
the fitted values needed by the RESET test, but it finds power up to
four is usually sufficient. Thus, we follow this guideline to conduct
the RESET test. For the Breusch-Pagan test, the choice of regressors in
the auxiliary regression is left to the user
\citep{breusch_simple_1979}. But as \citet{waldman1983note} suggested,
it is a good choice for the set of auxiliary regressors in the
Breusch-Pagan test be the same as the White test. Thus, we include both
\(\boldsymbol{x}\) and \(\boldsymbol{x}^2\) in the auxiliary regression.
